---
title: "JaB: Detecting Influential Nodes in Chilean Power Grid"
output: html_notebook
---

```{r setup, include = F}
set.seed(12)

## Libraries --------------------------------------------------------
# Needed for calculations
library(tidyverse)
library(igraph)
library(snowboot)

#Neeed to read data 
library(readxl)

# Needed for plotting
library(viridis)
library(ggrepel)
library(GGally)
invisible(lapply(c("ggplot2", "maps", "network", "sna"), base::library, character.only = TRUE))


## Functions --------------------------------------------------------
source("funcs/JaB-functions.R")



```


## The Data

We replicate the Chilean power grid results presented in Olivia Beck's Comprehensive Exam writen component. 



### Make an igraph object 

```{r}
### Data
############################
Chile_edgelist_orig <- read.csv("data/WithTap_edge.csv")
Chile_attributes <-  read.csv("data/WithTap_node.csv")

unique(Chile_edgelist_orig$Type) # all undirected 

#make igraph object 
edgelist <- as.matrix(Chile_edgelist_orig[, 1:2])
chile_igraph <- graph_from_edgelist(edgelist, directed = FALSE)

#plot it just to look
latlon <- as.matrix(Chile_attributes[, 5:6])
plot(chile_igraph, vertex.size = 1, layout = latlon, vertex.label = NA)


## Characteristics 
gorder(chile_igraph) #nodes 
length(E(chile_igraph)) #edges 


### Plot of original centarlity stats
data.frame(stat = log(igraph::betweenness(chile_igraph)+1)) %>%
  ggplot(aes(x = stat, y = ..density..)) + 
  geom_histogram(fill = "gray", bins = 19) +
  geom_density(adjust = 0.30) +
  theme_minimal()+
  xlab("")


```

### Running the Simulation 

We ran the following simulation in ROAR, Penn State's high-performance computing cluster. The results are saved the in the data/Chile-sim-results folder of this repository. 

```{}
### Options -------------------------------------------------
B.opts <- 1000
U.opts <- 100
s.opts <- c(1,2)
w.opts <- c(1,2,3)
q.opts <- c(0.9, 0.95, 0.99)
mult.opts <- c(TRUE, FALSE)
stat.opts <- c("betweenness", "eigen")

opts <- expand.grid(B.opts, U.opts, s.opts, w.opts, q.opts, mult.opts, stat.opts)
colnames(opts) <- c("B", "U", "s", "w", "q", "mult", "stat")

#save(opts, file = "rao-poster/sim-results/opts.Rdata")

### setting up the experiment
# only need to keep results matrix, dont need to keep null distributsion 

for( i in 1:nrow(opts)){
  
  #initalize storage and erase old storage 
  results.temp <- NA
  results <- NA
  
  #run algorithm with results
  results.temp <- null_dist_algorithm2(
    chile_igraph,  
    B=opts$B[i] , 
    U=opts$U[i],
    boot_type = "snowboot", 
    num_seed=opts$s[i], 
    num_wave=opts$w[i], 
    centrality_stat = opts$stat[i], 
    quants = c(0, opts$q[i]),
    mult = TRUE)
  
  #save 
  results <- results.temp$Results
  save(results, file = paste0("rao-poster/sim-results/results-opts-", i, ".Rdata"))
  
  
}
```


### Analyzing the Restuls 


```{r}
#save options from simiulation 
B.opts <- 1000
U.opts <- 100
s.opts <- c(1,2)
w.opts <- c(1,2,3)
q.opts <- c(0.9, 0.95, 0.99)
mult.opts <- c(TRUE, FALSE)
stat.opts <- c("betweenness", "eigen")

opts <- expand.grid(B.opts, U.opts, s.opts, w.opts, q.opts, mult.opts, stat.opts)
colnames(opts) <- c("B", "U", "s", "w", "q", "mult", "stat")


#extract the .Rdata objects saved from the ROAR simulation
file.names <- list.files("data/Chile-sim-results/", "results*")

load(paste0("data/Chile-sim-results/", file.names[1]))
keep.results <- data.frame(results, opts = 1)

#get all the nodes  flagged and sort into appropriate list 
for(i in 2:nrow(opts)){
  #load in new one
  load(paste0("data/Chile-sim-results/", file.names[i]))
  
  
  #get the file number #NOT the ordering of opts
  temp <- strsplit(file.names[i], split ="-")[[1]][3]
  temp2 <- strsplit(temp, split = "\\.")[[1]][1]
  
  #bind to old ones 
  keep.results <- rbind(keep.results, data.frame(results, opts = temp2 ))
}


keep.results <- as_tibble(keep.results)

```



```{r}
### Extract nodes that were flagged under betweenness for any q and how often
## constants for exploring 
opts90 <- which(opts$q == 0.90)
opts95 <- which(opts$q == 0.95)
opts99 <- which(opts$q == 0.99)

optsb <- which(opts$stat == "betweenness")

## q = 90, betweenness 
results90b <- keep.results %>% filter(opts %in% opts90) %>% filter(opts %in% optsb)

tab90 <- 
  results90b %>%
  filter(Influential == TRUE) %>%
  select(Node_Number) %>%
  table()

names(tab90) <- Chile_attributes$Label[as.numeric(names(tab90))]
tab90

## q = 95, betweenness 
results95b <- keep.results %>% filter(opts %in% opts95) %>% filter(opts %in% optsb)

tab95 <- results95b %>%
  filter(Influential == TRUE) %>%
  select(Node_Number) %>%
  table() 

names(tab95) <- Chile_attributes$Label[as.numeric(names(tab95))]
tab95

## q = 99, betweenness 
results99b <- keep.results %>% filter(opts %in% opts99) %>% filter(opts %in% optsb)

tab99 <- results99b %>%
  filter(Influential == TRUE) %>%
  select(Node_Number) %>%
  table() 

names(tab99) <- Chile_attributes$Label[as.numeric(names(tab99))]
tab99

```





### Ploting the maps 

```{r}
#make a network object 
chile.net <- network(Chile_edgelist_orig, directed = F)

# add coordniates to network
chile.net %v% "lat" <- Chile_attributes[network.vertex.names(chile.net), "Latitude"]
chile.net %v% "lon" <- Chile_attributes[network.vertex.names(chile.net), "Longitude"]


## Base plot with no zooming ------------------------------------------------------
chile <- 
  ggplot(map_data("world", region = "Chile"), aes(x = long, y = lat)) +
    geom_polygon(aes(group = group), color = "grey65",
                      fill = "grey", linewidth = 0.2)+
    theme_minimal()
chile


## Add network to plot -------------------------------------------
node.c <- rep("red", 347)
node.c[47] <- "blue"

node.s <- rep(0.25, 347)
node.s[47] <- 4

p_ <- GGally::print_if_interactive
p_ <- GGally::ggnetworkmap(
  chile, chile.net, size = node.s, great.circles = FALSE,
  segment.color = "black", node.color = node.c
)+ 
  xlim(c(-76, -66)) +
  ylim(c(-45, -17.5))

p_

### full map on smaller areas ---------------------------------------------

lat.min.0 <- -45
lat.max.0 <- -24
lon.min.0 <- -75
lon.max.0 <- -68

chile.zoom0.data <- map_data("world", region = "Chile") %>%
  dplyr::filter(lat < lat.max.0 & lat >lat.min.0 ,
                long < lon.max.0 & long > lon.min.0)

chile.zoom1 <- 
  ggplot(chile.zoom0.data, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), color = "grey65",
               fill = "grey", linewidth = 0.2)+
  theme_minimal()
chile.zoom1

## Add network to plot 
keep.nodes <- Chile_attributes %>%
  dplyr::filter(Latitude < lat.max.0 & Latitude > lat.min.0, 
                Longitude < lon.max.0 & Longitude > lon.min.0)

keep.rows <- Chile_edgelist_orig$Source %in% keep.nodes$Id & Chile_edgelist_orig$Target %in% keep.nodes$Id

keep.edges <- Chile_edgelist_orig[keep.rows, ]

chile.zoom0.net <- network(keep.edges, directed = F)

chile.zoom0.net %v% "lat" <- Chile_attributes[network.vertex.names(chile.zoom0.net), "Latitude"]
chile.zoom0.net %v% "lon" <- Chile_attributes[network.vertex.names(chile.zoom0.net), "Longitude"]

color.vec0 <- rep("#4e51bf", length(network.vertex.names(chile.zoom0.net)))


p_0 <- GGally::print_if_interactive
p_0 <- GGally::ggnetworkmap(
  chile.zoom1, chile.zoom0.net, size = 0.5, great.circles = FALSE,
  segment.color = "black", node.fill = color.vec0, node.color = color.vec0
)

#color data 
dat.extra <- data.frame(Id = c(105, 108, 264, 178, 130),
                        lat = NA,
                        long = NA)
for(i in 1:5){
  dat.extra$lat[i] <- Chile_attributes$Latitude[Chile_attributes$Id == dat.extra$Id[i]]
  dat.extra$long[i] <- Chile_attributes$Longitude[Chile_attributes$Id == dat.extra$Id[i]]
}

#plot

p_0+ 
  xlim(c(lon.min.0, lon.max.0)) +
  ylim(c(lat.min.0, lat.max.0)) +
  geom_point(data = dat.extra, color = c("#c20320"), shape = 15)


### Zoomed in map on Cluster 1 (northern cluster) --------------------------------------------------------------

lat.min.1 <- -35.0
lat.max.1 <- -32.5
lon.min.1 <- -72.25
lon.max.1 <- -70

chile.zoom1.data <- map_data("world", region = "Chile") %>%
  dplyr::filter(lat < lat.max.1 & lat >lat.min.1 ,
                long < lon.max.1 & long > lon.min.1)

chile.zoom1 <- 
  ggplot(chile.zoom1.data, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), color = "grey65",
               fill = "grey", linewidth = 0.2)+
theme_minimal()
chile.zoom1

## Add network to plot 


keep.nodes <- Chile_attributes %>%
  dplyr::filter(Latitude < lat.max.1 & Latitude > lat.min.1, 
                Longitude < lon.max.1 & Longitude > lon.min.1)

keep.rows <- Chile_edgelist_orig$Source %in% keep.nodes$Id & Chile_edgelist_orig$Target %in% keep.nodes$Id

keep.edges <- Chile_edgelist_orig[keep.rows, ]

chile.zoom1.net <- network(keep.edges, directed = F)

chile.zoom1.net %v% "lat" <- Chile_attributes[network.vertex.names(chile.zoom1.net), "Latitude"]
chile.zoom1.net %v% "lon" <- Chile_attributes[network.vertex.names(chile.zoom1.net), "Longitude"]



p_1 <- GGally::print_if_interactive
p_1 <- ggnetworkmap(
  chile.zoom1, chile.zoom1.net, size = 3, great.circles = FALSE,
  segment.color = "black", node.color = "#4e51bf"
)

#color data 
dat.extra0 <- data.frame(Id = c(105, 130),
                        lat = NA,
                        long = NA)
for(i in 1:2){
  dat.extra0$lat[i] <- Chile_attributes$Latitude[Chile_attributes$Id == dat.extra0$Id[i]]
  dat.extra0$long[i] <- Chile_attributes$Longitude[Chile_attributes$Id == dat.extra0$Id[i]]
}

#plot
p_1+ 
  xlim(c(lon.min.1, lon.max.1)) +
  ylim(c(lat.min.1, lat.max.1)) +
  geom_point(data = dat.extra, shape = 15, color = "#c20320", size = 6)


### Zoomed in map on Cluster 2 (southern cluster) --------------------------------------------------------------

lat.min.2 <- -38.25
lat.max.2 <- -36.5
lon.min.2 <- -74
lon.max.2 <- -70.5

chile.zoom2.data <- map_data("world", region = "Chile") %>%
  dplyr::filter(lat < lat.max.2 & lat >lat.min.2 ,
                long < lon.max.2 & long > lon.min.2)

chile.zoom2 <- 
  ggplot(chile.zoom2.data, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), color = "grey65",
               fill = "grey", linewidth = 0.2)+
  theme_minimal()
chile.zoom2

## Add network to plot


keep.nodes <- Chile_attributes %>%
  dplyr::filter(Latitude < lat.max.2 & Latitude > lat.min.2, 
                Longitude < lon.max.2 & Longitude > lon.min.2)

keep.rows <- Chile_edgelist_orig$Source %in% keep.nodes$Id & Chile_edgelist_orig$Target %in% keep.nodes$Id

keep.edges <- Chile_edgelist_orig[keep.rows, ]

chile.zoom2.net <- network(keep.edges, directed = F)

chile.zoom2.net %v% "lat" <- Chile_attributes[network.vertex.names(chile.zoom2.net), "Latitude"]
chile.zoom2.net %v% "lon" <- Chile_attributes[network.vertex.names(chile.zoom2.net), "Longitude"]



p_2 <- GGally::print_if_interactive
p_2 <- ggnetworkmap(
  chile.zoom2, chile.zoom2.net, size = 3, great.circles = FALSE,
  segment.color = "black", node.color = "#4e51bf"
)


#plot
p_2+ 
  xlim(c(lon.min.2, lon.max.2)) +
  ylim(c(lat.min.2, lat.max.2)) +
  geom_point(data = dat.extra, shape = 15, color = "#c20320", size = 6)

### Zoomed in map on Cluster 3 (middle cluster) --------------------------------------------------------------

lat.min.3 <- -36.5
lat.max.3 <- -34.6
lon.min.3 <- -73
lon.max.3 <- -70

chile.zoom3.data <- map_data("world", region = "Chile") %>%
  dplyr::filter(lat < lat.max.3 & lat >lat.min.3 ,
                long < lon.max.3 & long > lon.min.3)

chile.zoom3 <- 
  ggplot(chile.zoom3.data, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), color = "grey65",
               fill = "grey", linewidth = 0.2)+
  theme_minimal()
chile.zoom3

## Add network to plot


keep.nodes <- Chile_attributes %>%
  dplyr::filter(Latitude < lat.max.3 & Latitude > lat.min.3, 
                Longitude < lon.max.3 & Longitude > lon.min.3)

keep.rows <- Chile_edgelist_orig$Source %in% keep.nodes$Id & Chile_edgelist_orig$Target %in% keep.nodes$Id

keep.edges <- Chile_edgelist_orig[keep.rows, ]

chile.zoom3.net <- network(keep.edges, directed = F)

chile.zoom3.net %v% "lat" <- Chile_attributes[network.vertex.names(chile.zoom3.net), "Latitude"]
chile.zoom3.net %v% "lon" <- Chile_attributes[network.vertex.names(chile.zoom3.net), "Longitude"]



p_3 <- GGally::print_if_interactive
p_3 <- ggnetworkmap(
  chile.zoom3, chile.zoom3.net, size = 3, great.circles = FALSE,
  segment.color = "black", node.color = "#4e51bf"
)


#plot
p_3+ 
  xlim(c(lon.min.3, lon.max.3)) +
  ylim(c(lat.min.3, lat.max.3)) +
  geom_point(data = dat.extra, shape = 15, color = "#c20320", size = 6)



```